Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex
correction 2

\section{Related work}
\label{sec:related}
Image segmentation and object recognition in medical imagining have 
been tackled for many years and many approaches have been found. 
Methods based on mathematical morphological operators have been 
widely used and applied for segmentation and feature extraction, 
however such methods rely on the user to manually define the 
information that is relevant for his task \cite{irshad2014methods}. 
Segmentation methods based on a strong priori information are also 
possible and yield comparable results in terms of accuracy, in 
\cite{ranefall2016fast} they segment the picture using a ellispe fit 
which is considerably faster than other state of the art methods. Graph-
based tools for segmentation are also possible in histopathology 
\cite{ta2009graph}. \cite{manivannanlocal} focus on gland 
segmentation, they achieve good results by predicting image patches as 
a whole, via a classification scheme. The output confidence vector is 
then preprocessed to obtain the final segmentation. Alternative methods 
have arisen from the achieving Convolutional neural networks 
\cite{lecun}, in particular strong incentive rose with the very impressive 
results that they achieved in the pascal voc challenge with 
\cite{ImageNet}. Recent advances in deep neural network, and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition.  Deep neural network have also been 
used for different task and have yielded very impressiv results, we can think of 
semantic segmentation problems, where \cite{long2015fcn} use "de-
convolution layers" and up-sampling in order to identify and precisely 
locate objects within a picture. Different architecture arising from 
different intuition are also possible and have been applied in this paper.

Talk about related work.

@@

%% Contribution as a dataset, small introduction
One of the main contributions of this paper is the now public available 
nuclei detection dataset within HE stainned histopathology images which 
can be found at website %\url{www.cbio.ensmp.fr:/pnaylor/download/dataset.zip} 
(not a true address). This annotated dataset provides images clustered by 
patients. Each patient has at least 3 annotated $512 \times 512$ HE 
stainned histopathology images with their associated ground truth. Each 
ground truth image is a $512 \times 512$ image where each pixel value above 
$0$ is considered as a pixel belonging to a nuclei. The differences in 
values of these pixels denote different nucleus, such an annotation can 
be used in several processing step which does or doesn't take into 
account clutered nuclei. See figure \ref{fig:annotation} for an example of 
three annotated images. This annotation was conducted via the help of 
the software ITK-snap, \cite{py06nimg} http://www.itksnap.org, and were annotated by the authors of the paper.

@@

%%Data set description 
In this data set we have annotated a considerous amount of cells, 
including normal breast cells (epithelial and lobule cells),  cancerous 
breast cells, fibroblasts,  endothelial cells, fat cells, macrophage cells and 
inflammatory cells (lymphocytes and plasmocytes). For the moment, cell types have not been seperated, which means that in this dataset every annotation corresponds to a global type: cell.

\begin{itemize}
\item Number of images: 33 \\
\item Number  of cells: 2754 \\
\item Maximum number of cells in one sample: 293 \\
\item Minimum number of cells in one sample: 5 \\
\item Mean number of cells: 83 \\
\item Standard deviation of number of cells: 63 \\
\end{itemize}

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores on the test sets.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
We conducted our experiments on GPUs via the caffe framework, \cite{jia2014caffe}. We present our results in table \ref{tab:res} where we averaged the metrics other all left out patients. We also added, for illustration purposes \ref{fig:prediction}, image prediction and associated probability maps to evaluate some differences between the classifiers. We notice that the simpler net, PangNet seems to have learnt simple rules such as color information. As we can not notice by the "holes" in the probability map when the interior of the cell is brighter. On the contrary, deeper nets as FCN and DeconvNet have learnt deeper features and seem capable of recognizing whole cells. .
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

@@

\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanB}}
%  \vspace{1.5cm}
  \centerline{PangNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}

%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetB}}
%  \vspace{1.5cm}
  \centerline{DeconvNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNB}}
%  \vspace{1.5cm}
  \centerline{FCN}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
\caption{Prediction via different classifiers of a random sample on the left out patient: 572123}
\label{fig:prediction}
%
\end{figure}

>>>>

\section{Related work}
\label{sec:related}

In \cite{Drelie08-298}, they have a partially annotated data set of histopathology images for nuclei segmentation.

Image segmentation and object recognition in medical imagining have 
been tackled for many years and many approaches have been found. 
Methods based on mathematical morphological operators have been 
widely used and applied for segmentation and feature extraction, 
however such methods rely on the user to manually define the 
information that is relevant for his task \cite{irshad2014methods}. 
Segmentation methods based on a strong priori information are also 
possible and yield comparable results in terms of accuracy, in 
\cite{ranefall2016fast} they segment the picture using mathematical 
morphology with an ellispe fit condition 
which is considerably faster than other state of the art methods. 
%Graph-
%based tools for segmentation are also possible in histopathology 
%\cite{ta2009graph}. 
%\cite{manivannanlocal} focus on gland 
%segmentation, they achieve good results by predicting image patches as 
%a whole, via a classification scheme. The output confidence vector is 
%then preprocessed to obtain the final segmentation. 
Alternative methods 
have arisen from the achieving Convolutional neural networks 
\cite{lecun}, in particular strong incentive rose with the very impressive 
results that they achieved in the pascal voc challenge with 
\cite{ImageNet}. Recent advances in deep neural network, and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition.  Deep neural network have also been 
used for different task and have yielded very impressiv results, we can think of 
semantic segmentation problems, where \cite{long2015fcn} use "de-
convolution layers" and up-sampling in order to identify and precisely 
locate objects within a picture. Different architecture arising from 
different intuition are also possible and have been applied in this paper.

@@

%% Contribution as a dataset, small introduction
One of the main contributions of this paper is the now public available 
nuclei detection dataset within HE stainned histopathology images which 
can be found at website \url{http://cbio.mines-paristech.fr/~pnaylor/Downloads/HECellSegmentation.zip} 
(not a true address). This annotated dataset provides images clustered by 
patients. Each patient has at least 3 annotated $512 \times 512$ HE 
stainned histopathology images with their associated ground truth. Each 
ground truth image is a $512 \times 512$ image where each pixel value above 
$0$ is considered as a pixel belonging to a nuclei. The differences in 
values of these pixels denote different nucleus, such an annotation can 
be used in several processing step which does or doesn't take into 
account clutered nuclei. See figure \ref{fig:annotation} for an example of 
three annotated images. This annotation was conducted via the help of 
the software ITK-snap, \cite{py06nimg} \url{http://www.itksnap.org}, and were annotated by the authors of the paper.

@@

%%Data set description 
%% I asked Marick's help for this paragraph!
In this data set we have annotated a considerous amount of cells,
including normal epithelial and myoepithelial breast cells (localized in
ducts and lobules), invasive carcinomatous cells, fibroblasts, endothelial
cells, adipocytes, macrophages and
inflammatory cells (lymphocytes and plasmocytes). For the moment, cell 
types have not been included, which means that in this dataset every annotation corresponds to a global type: cell. We can also give some simple statistics with respect to the dataset. We have 33 images with a total of 2754 annotated cells, the maximum number of cells in one sample is 293 and the minimum number of cells in one sample is 5. We also have on average 83 cells per sample with a high standard deviation of 63.

% \begin{itemize}
% \item Number of images: 33 \\
% \item Number  of cells: 2754 \\
% \item Maximum number of cells in one sample: 293 \\
% \item Minimum number of cells in one sample: 5 \\
% \item Mean number of cells: 83 \\
% \item Standard deviation of number of cells: 63 \\
% \end{itemize}

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores on the test sets.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics other all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As we can not notice by the "holes" in 
the probability map when the interior of the cell is very bright. On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. Our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation or by simple 
visualation of the results.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

\section{Conclusion}
\label{sec:conclusion}

%I don't really know if we need something like this...

@@

\begin{minipage}[b]{0.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{BaochuanB}}
%  \vspace{1.5cm}
  \centerline{PangNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.5cm]{BaochuanP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{Baochuan_WS}}
%  \vspace{1.5cm}
  \centerline{Watershed}\medskip
\end{minipage}

%
%
\begin{minipage}[b]{0.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{DeconvNetB}}
%  \vspace{1.5cm}
  \centerline{DeconvNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{DeconvNetP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{DeconvNet_WS}}
%  \vspace{1.5cm}
  \centerline{Watershed}\medskip
\end{minipage}
%
\begin{minipage}[b]{0.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{FCNB}}
%  \vspace{1.5cm}
  \centerline{FCN}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{FCNP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.32\linewidth}
  \centering
  \centerline{\includegraphics[height=2.4cm]{FCN_WS}}
%  \vspace{1.5cm}
  \centerline{Watershed}\medskip
\end{minipage}
%
\caption{Prediction via different classifiers of a random sample on the left out patient: 581910}
\label{fig:prediction}
%
\end{figure}
